Recipient plant traits are important determinants of the impacts of heterospecific pollen upon plant reproduction

### 2019-03-18 Manuscript_draft/Text.Rmd
trying to add more results

Results of hand cross-pollination, hand self-pollination, natural selfing and apomixis are presented for all the species in **Figure 1** (see **Appendix 1** for table with values). Heterospecific pollen reduced seed set significatively with the 50-50% heterospecific pollen treatments for 65% of the pairwise interactions p<0.05. Brassicaceae and Solanaceae species had less foreign pollen than expected, the ratios of pollen of recipient and donor respectively for both families were 80-20% and 60.5-39.5%. In contrast, Convolvulaceae species had just 28% recipient pollen and 72% of donor pollen. In addition, for the 100% hetrospecific pollen treatments we did not find almost seed production. However, for just one species (*S. alba*) the control pollination and the heterospecific pollination with pollen from a confamilial had similar seed production. For two Solanaceae species *S. melongena* and *C. annuum* 100% pollen treatments produced few seedles fruits (3% and 9% respectively) and they did not for the apomictic treatments.

The grouped effect size of the nine donors per species differed across species and families (see **Figure 2**). Interestingly, despite some variability in the effect of the different pollen donors per species, in general terms the effect of heterospecific pollen from the distinct nine treatments per species was homogeneous (see **Figure 3**). Just for four species out of ten, just one donor in each did not overlap the confidence intervals with other donors. Therefore, none of the donors had a clear stronger or weaker effect across species. When the donors were grouped by family we found a similar effect too, just for *S. lycopersicum* the confidence intervals of Brassicaceae and Convolvulaceae did not overlap (see **Figure 4**).

[add numeric values for 100% seed production](Jose)

Results of Mantel test between heterospecific pollen effect and phylogenetic distance gave a positive statistically clear correlation for both markers (p<0.05). The correlations with ITS and RBCL markers was respectively of 0.29 and 0.25. We found a significant phylogenetic signal of traits for pollen size, stigma measurements and style length (p<0.05). Although with a lack of a significant correlation pagel's lambda values were also relatively high (>0.45) for incompatibility index, ovary length and levels of selfing.

Moreover, Mantel test between heterospecific pollen effect and the distance matrix of traits gave also a positive significant correlation with r value of 0.4. When the effect was look trait by trait with Mantel test, stigma type and stigma measurements (length, width and area) gave a significant positive correlation with heterospecific pollen effect. When the effect of heterospecific pollen was just looked at family level, for Solanaceae species we found a negative correlation between style length and heterospecific pollen effect (Figure 7).

>>>>

Results of hand cross-pollination, hand self-pollination, natural selfing and apomixis are presented for all the species in **Figure 1** (see **Appendix 1** for table with values). Heterospecific pollen reduced seed set significatively with the 50-50% heterospecific pollen treatments for 65% of the pairwise interactions p<0.05. Brassicaceae and Solanaceae species had less foreign pollen than expected, the ratios of pollen of recipient and donor respectively for both families were 80-20% and 60.5-39.5%. In contrast, Convolvulaceae species had just 28% recipient pollen and 72% of donor pollen. In addition, for the 100% hetrospecific pollen treatments we did not find almost seed production (see *Table 3* for 100% treatments that produced seeds). However, for just one species (*S. alba*) the control pollination and the heterospecific pollination with pollen from a confamilial had similar seed production (non significant differences). Moreover, for two Solanaceae species *S. melongena* and *C. annuum* 100% pollen treatments produced few seedles fruits (3% and 9% respectively) and they did not for the apomictic treatments indicating that foreing pollen can be responsible of seedless fruit production.

The grouped effect sizes of the nine donors per species differed across species and families (see **Figure 2**), having clear statistical differences for species within and across families with not overlapping confidence intervals. Interestingly, despite some variability in the effect of the different pollen donors per species, in general terms the effect of heterospecific pollen from the distinct nine treatments per species was homogeneous with unclear statistical differences (see **Figure 3**). Just for four species out of ten, just one donor in each did not have overlapping confidence intervals with the rest of the donors. Therefore, none of the donors had a clear stronger or weaker effect across species. When the donors were grouped by family we found a similar effect too, just for *S. lycopersicum* the confidence intervals of Brassicaceae and Convolvulaceae did not overlap (see **Figure 4**).

Matrices of heterospecifc pollen effect and traits (Distance matrix of traits, euclidean distance) are positively and significantly correlated (r=0.4, one-tailed Mantel test p<0.05).The variable with maximum rank correlation (thorugh *Bioenv* funtion) was incompatibility index by itself within a group of 16 variables calculated with euclidean distances.

[Add Bioenv results](Jose)

The same analysis of Mantel test between HP pollen effect and the traits one by one gave a positive and significant correlation for stigma type and stigma measurements (length, width and area). When the effect of heterospecific pollen was looked at family level, for Solanaceae species we found a strong negative non significant correlation between style length and heterospecific pollen effect (Pearson's correlation=0.93, p=0.07) (see **Figure 7**).

Results of Mantel test between heterospecific pollen effect and phylogenetic distance gave a low positive correlation for both markers (p<0.05). The correlations with ITS and RBCL markers was respectively of 0.29 and 0.25. We found a significant phylogenetic signal of traits for pollen size, stigma measurements and style length (p<0.05). Although with a lack of a significant correlation pagel's lambda values were also relatively high (>0.45) for incompatibility index, ovary length and levels of selfing.
